\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-03-24 response.tex

2. Proof of Theorem 9: There is too much notation throughout. As
one example, it is hard to remember which of $T_x$, $Y_x$, $V_x$, $F_x$, $N_x$,
$S_x$, $X_v$, $B_x$, $C_x$ is which. If $\phi(S)$ is written for the subset of $V(G)$
corresponding to a set $S \subseteq V(H)$ and appropriate shorthand and
standard notation for neighbourhoods are used, then $Y_x$, $V_x$, $F_x$, and
$N_x$ can be replaced by $\phi(x)$, $\phi(T_x)$, $N(\phi(T_x))$, and $V(G) \setminus N[\phi(T_x)]$.
The set $X_v$ is only used in the statement of Claim~1, so it does not
need a name; say what these vertices are explicitly. It may be more
indicative to write $\phi'(x)$ for $S_x$ and $B'_x$ for $C_x$.

\begin{response}
  We agree that the notation was excessive (an unfortunate consequence of the general result on $(k,d)$-shortcut systems having evolved from a specific result about $k$-planar graphs).  We have eliminated the notations $V_x$, $N_x$, $F_x$, $X_x$, and $T_x$.

  The notations $Y_x$ and $B_x$ are needed since we need names for the parts of the $H$-decomposition of $G$ and the bags in the tree-decomposition of $H$.    The notation $C_x$ is limited to the proof of (newly-numbered) Claim~3, where it is used to denote the bags in the tree-decomposition that is the subject of Claim~3.

  We have avoided the introduction of any new notation, including $\phi$.  The new presentation is much shorter and clearer.  We thank the referee for making us revisit this proof carefully.  The final paper is better for it.
\end{response}

@@

3. Theorem 10: Unfortunately I have to suggest removing it. I don’t see
the motivation for making $H$ planar or for improving the constants.
Are the improved constants at all close to tight? Is there another
reason for including it?

\begin{response}
  Since the initial submission, there is now even more justification for including this special case.  A result of Bose et al on vertex ranking relies critically on the simple treewidth of $H$, as does the result of D\,ebski et al on $p$-centered colouring.  In both case, if $H$ does not have simple treewidth $3$ then one does not get an asymptotically tight bound.

  Also, while adding details requested by Referee \#2 we discovered that a small change in the proof (removing both edges of a crossing pair instead of just one) results in a better constant and leads to a more general result that applies to $d$-framed graphs. This, in turn, leads to better results for $d$-map graphs.

  The former Theorem~10 (on $1$-planar graphs) is now a consequence of  Theorem~23 on ($d$-framed graphs).
\end{response}

@@

$P$ is used twice in the statement of Theorem 3.  It is probably
safer (and necessary?) to use a different variable for the second
occurrence.

\textcolor{blue}{We now use $\mathcal{S}$ for shortcut systems.}

Page 6. Regarding (T1), technically a subtree of a rooted tree can be
rooted at any vertex, so it is a bit imprecise to say that $T[x]$ is
rooted at $x$.  I suggest adding that a subtree of a tree rooted at $r$,
is always considered to be rooted at the vertex closest to $r$.

\textcolor{blue}{Fixed, though in a different way.  Since we have to define lowest common $T$-ancestors anyway, we define the subtree $T'$ of $T$ to be rooted at at the lowest common $T$-ancestor of $V(T')$.}

### 2022-03-24 response.tex

  The notations $Y_x$ and $B_x$ are needed since we need names for the parts of the $H$-decomposition of $G$ and the bags in the tree-decomposition of $H$.    The notation $C_x$ is limited to the proof of (newly-numbered) Claim~3, where it is used to denote the bags in the tree-decomposition that is the subject of Claim~3.

  We have avoided the introduction of any new notation, including $\phi$.  The new presentation is much shorter and clearer.  We thank the referee for making us revisit this proof carefully.  The final paper is better for it.
\end{response}

3. Theorem 10: Unfortunately I have to suggest removing it. I don’t see
the motivation for making $H$ planar or for improving the constants.
Are the improved constants at all close to tight? Is there another
reason for including it?

\begin{response}
There is a lot of motivation for proving this result with $H$ of treewidth 3 and planar. First, the treewidth bound is best possible (even for planar graphs). Having $H$ planar means that $H$ is in fact a planar 3-tree (also called stack polytope or Appolonian etwork). It follows that we get better bounds for queue-number for example.

Since the initial submission, there is now even more justification for including this material.  A result of Bose et al on vertex ranking relies critically on the simple treewidth of $H$, as does the result of Debski et al on $p$-centered colouring.  In both case, if $H$ does not have simple treewidth $3$ then one does not get an asymptotically tight bound.

### 2022-03-24 k-planar.tex
comments and revisisions

\begin{table}[h]
  \centering{
    \begin{tabular}{lccl}
      \textbf{Graph class} & \boldmath $\tw(H)$ & \boldmath $\ell$ & \textbf{Reference} \\ \hline
      planar & $3^*$ & $3$ & \cite{DJMMUW20} \\
      planar & $6^*$ & $1$ & \cite{UWY} \\
      genus $g$ & $3^*$ & $\max\{2g,3\}$ & \cite{DHHW} \\
%      genus $g$ & $2g+6$ & 1 & \cite{UWY} \\
      apex-minor-free & $O(1)$ & $O(1)$ & \cite{DJMMUW20} \\[1ex]
      % B-minor-free & $O(1)$ & $O(1)$ & \cite{DJMMUW20} \\
      $k$-planar & $\binom{k+4}{3}-1$ & $18k^2 + 48k + 30$ & \cref{kPlanarProduct} \\
      $(g,k)$-planar & $\binom{k+4}{3}-1$ & $\max\{2g,3\}(6k^2+16k+10)$ & \cref{gkPlanarProduct} \\
      $(g,\delta)$-string & $\binom{\delta+4}{3}-1$ & $\max\{2g,3\}(\delta^4 + 4\delta^3 + 9\delta^2 + 10\delta+4)$ & \cref{StringProduct} \\
      $k$-NN & $O(k^6)$ & $O(k^4)$ & \cref{nn_product_structure} \\
      $d$-framed & $3^*$ & $d+ 3\floor{d/2} - 3$ & \cref{d_framed_product_stucture} \\
      $1$-planar & $3^*$ & $7$ & \cref{1_planar_product} \\
      $d$-map & $3^*$ & $d+ 3\floor{d/2} - 3$ & \cref{dMapProduct} \\
      $(g,d)$-map & $9$ & $\max\{2g,3\}\,(7d^2 -21d)$ & \cref{gdMapProduct} \\ \hline 
      \multicolumn{4}{l}{$^*$ these bounds also applies to the simple treewidth of $H$.}
    \end{tabular}
  }
  \caption{Known product structure theorems (of the form $G\subseteq H\boxtimes P\boxtimes K_\ell$).}
  \label{summary_table}
\end{table}

@@

@@

Every $d$-framed multigraph can be described by a $(2,d(d-3)/2)$-shortcut system applied to a plane multigraph by adding a vertex inside each face $F$ of $G_0$ adjacent to the vertices of $F$ and creating a shortcut between each non-consecutive pair of vertices of $F$. Thus \cref{ShortcutProduct} is applicable. However, for framed graphs we prove the following stronger result, where the treewidth bound on $H$ is best possible.\footnote{A multigraph $G$ is \defin{contained} in a graph $G'$ if the simple graph underlying $G$ is isomorphic to a subgraph of $G'$.}

@@

	The proof is by induction on the number of inner faces of $N_0$.  A particular easy case occurs when $N_0$ has no inner vertices, i.e., $V(N_0)=V(F)$. In this case we can simply take
	\note{DW}{Suggested revision, ``In the base case, $N_0$ has only one inner face, and we can simply take ...} 
  % \note{PM}{I debated what to put here since, even when $N'$ has only one face, the rules below work for picking $\tau$.  The \emph{real} base case occurs when $N'$ has zero faces, but that's not something that ever happens since we only ever apply induction on the inner faces of $M$ (except $\tau$).  I settled on this ``base case'' since it's clear that, in this case, the partition $\mathcal{P}$ is already decided.}
  \note{PM}{The longwinded version is that induction is on $(x,y)$ where $x$ is the number of inner vertices of $N_0$ and $y$ is the number of inner faces of $N_0$.  The base cases are $(0,y)$ for any $y$.  Unfortunately, we can't just induct on $x$ because it doesn't always decrease.}
	$\PP:=\{P_1,\ldots,P_k\}$ and verify that the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements (i)--(iii):
	\begin{compactenum}[(i)]
		\item By definition, each of $P_1,\ldots,P_k$ is a part of $\mathcal{P}$.
		\item Assumption 4(b) ensures, for each $x\in V(H)$, the existence of $X_x$ and $Y_x$ such that $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each integer $j\ge 0$.
		\item The trivial tree decomposition of $H$ that has a single bag containing $\{x_1,\ldots,x_k\}$ is $3$-simple and, if $k=3$, has exactly one bag that contains $\{x_1,\ldots,x_3\}$.
	\end{compactenum}

@@

\begin{proof}
% \note{PM}{I fixed some issues here that come from the fact that some vertices of $G_0$ do not contribute faces to $G_1$.  This happens when two nations and two lakes meet at a vertex $x$.  It's not an issue but it means we can't write ``the face of $G_1$ that corresponds to $x$.''}
Let $G_0$ be a graph embedded in the plane, with each face labelled a nation or a face, and where each vertex of $G_0$ is incident with at most $d$ nations. If two lakes have a common edge in their boundaries, then we may delete the edges and merge the lakes. Now assume that no two lakes have a common edge in their boundaries.
Let $G$ be the corresponding map graph.
Let $G_1$ be the graph with $V(G_1):=V(G)$, which is the set of nations, and whose edge set we now define.  Consider vertices $v$ and $w$ of $G_1$.
Let $A$ and $B$ be the nation faces of $G_0$ corresponding to $v$ and $w$.
Then $vw\in E(G_1)$ if $A$ and $B$ have a common edge in their boundaries,
or there is a vertex $x$ of $G_0$ in the boundary of $A$ and $B$, and in the clockwise or anticlockwise order of faces around $x$, all the faces between $A$ and $B$ are lakes.
% Then $G_1$ has Euler genus at most $g$.
Each face of $G_1$ corresponds to a vertex of $G_0$ or a lake of $G_0$.
Triangulate each face of $G_1$ corresponding to a lake of $G_0$.

For each face $F$ of $G_1$ that corresponds to a vertex $x$ of $G_0$, the size of $F$ equals the number of nations incident to $x$, which is at most $d$.
So every face of $G_1$ has size at most $d$.  Let $\widehat{G}_1$ be the $d$-framed graph whose frame is $G_1$.

To complete the proof, we now show that $\widehat{G}_1$ contains every edge of $G$.
If $vw\in E(G)$ then the nation faces of $G_0$ corresponding to $v$ and $w$ have a vertex $x$ of $G_0$ in common.   If some face of $G_1$ corresponds to $x$ then $v$ and $w$ are on a common face of $G_1$ so $vw\in E(\widehat{G}_1)$.  If no face of $G_1$ corresponds to $x$\footnote{This can occur when $x$ is on the boundary of exactly two nation faces and two lake faces.} then the only two nation faces of $G_0$ with $x$ on their boundary are those corresponding to $v$ and $w$, in which case $vw\in E(G_1)\subseteq E(\widehat{G}_1)$.  Therefore $G\subseteq\widehat{G}_1$.
\end{proof}

@@

As discussed in the introduction, product structure has been used to resolve a number of problems on planar graphs.  In most cases, these results hold for any graph with product structure. In this section, we survey some of the consequences of this for the graph classes considered in the previous two sections.

@@

For an integer $k\geq 0$, a \textit{$k$-queue layout} of a graph $G$ consists of a linear ordering $\preceq$ of $V(G)$ and a partition $\{E_1,E_2,\dots,E_k\}$ of $E(G)$, such that for $i\in\{1,2,\dots,k\}$, no two edges in $E_i$ are nested with respect to $\preceq$. That is, it is not the case that $v\prec x \prec y \prec w$ for edges $vw,xy\in E_i$. The \textit{queue-number} of a graph $G$, denoted by $\qn(G)$, is the minimum integer $k$ such that $G$ has a $k$-queue layout. Queue-number was introduced by \citet{HLR92}, who famously conjectured that planar graphs have bounded queue-number. \citet{DJMMUW20} recently proved this conjecture using \cref{PlanarProduct} and the following lemma. Indeed, resolving this question was the motivation for the development of \cref{PlanarProduct}.

@@

Since all the product structure theorems presented thus far upper bound the treewidth of $H$,
\cref{qn} immediately implies bounds on the queue number of all graphs in these classes.  An improvement can be made for $1$-planar and $d$-map graphs since, in the cases $H$ is planar and has treewidth $3$.
\citet{ABGKP20} proved that every planar graph with treewidth at most $3$ has queue-number at most $5$. Thus the graph $H$ in \cref{1_planar_product,dMapProduct} has queue-number at most $5$.
% , which with \cref{qn} implies:
% 
% \begin{cor}
% \label{1PlanarQueue}
% Every 1-planar graph has queue-number at most 
% $3 \times 7 \times 5 + \floor{\tfrac{3}{2} \times 7} = 115$.
% \end{cor}
% 
% \begin{cor}
% \label{dMapQueue}
% Every $d$-map graph has queue-number at most 
% $\floor{ \frac{33}{2} (d+3\floor{d/2}-3) }$.
% \note{DW}{I have updated this bound using the better qn bound for planar 3-trees. Note that with $d=3$ we get the original bound of 49.}
% \end{cor}
% 
% 
% 
% \cref{qn,gkPlanarProduct} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^3)}$.
% \note{DW}{I have changed $g 2^{O(k^4)}$ to $g 2^{O(k^3)}$, okay?}
% \note{OM}{Ok}
% \note{DW}{I have changed $g>2^{k^3}$ to $g>2^{k^2}$, okay?}.
% \note{PM}{Ok}
The following corollary summarizes all of these results.
\begin{cor}\label{q_cor}
  The following bounds hold for the queue number of any graph from each of the following classes:
  \begin{compactenum}
    \item $k$-planar: $2^{O(k^3)}$
    \item $(g,k)$-planar: $\max\{1,g\}\cdot 2^{O(k^3)}$
    \item $(g,\delta)$-string: $\max\{1,g\}\cdot 2^{O(\delta^3)}$
    \item $1$-planar: $115$ 
    \item $d$-map: $\floor{ \frac{33}{2} (d+3\floor{d/2}-3) }$ 
    \item $(g,d)$-map: $\max\{1,g\}\cdot 2^{O(d)}$
    \item $k$-NN: $2^{O(k^6)}$
    % \item $(g,d)$-map graph
    % \item
  \end{compactenum}
\end{cor}
\note{PM}{For uniformity, I got rid of the two corollaries on $d$-map and $1$-planar graphs and kept them in \cref{q_cor}.  The text is still commented out in the source in case someone feels strongly about keeping them.}

@@

Combining \cref{non-repetitive} with our results on product structure, we obtain the following corollary:
\begin{cor}\label{nonrepetitive_cor}
  The following bounds hold for the nonrepetitive chromatic number of any graph from each of the following classes:
  \begin{compactitem}
    \item $k$-planar: $(18k^2+48k+30)\cdot 4^{\binom{k+4}{3}}$
    \item $(g,k)$-planar: $\max\{2g,3\}\cdot(6k^2+16k+10) 4^{\binom{k+4}{3}}$
    % \item $(g,\delta)$-string: $g4^{O(\delta^3)}$
    \item $1$-planar: $1792$
    \item $d$-map: $256(d+3\floor{d/2}-3)$
    % \item $(g,d)$-map: $g4^{O(d)}$
    % \item $k$-NN: $4^{O(k^6)}$
    \item $(g,d)$-map graph: $(7d^2-21d)\cdot 4^{10}$
  \end{compactitem}
\end{cor}

@@

\begin{cor}\label{p_centered_cor}
  The following bounds hold for the $p$-centered chromatic number of any graph from each of the following classes:
  \begin{compactitem}
    \item $k$-planar: $(18k^2+48k+30)\cdot 4^{\binom{k+4}{3}}$
    \item $(g,k)$-planar: $\max\{2g,3\}\cdot(6k^2+16k+10)(p+1)\binom{p+\binom{k+4}{3}-1}{\binom{k+4}{3}-1}$
    % \item $(g,\delta)$-string: $g4^{O(\delta^3)}$
    \item $1$-planar: $7(p+1)^2(p\ceil{\log(p+1)}+2p+1)$
    \item $d$-map: $(d+3\floor{d/2}-3)\cdot (p+1)^2(p\ceil{\log(p+1)}+2p+1)$
    % \item $(g,d)$-map: $g4^{O(d)}$
    % \item $k$-NN: $4^{O(k^6)}$
    \item $(g,d)$-map: $\max\{2g,3\}\cdot 7d(d-3)(p+1)\cdot \binom{p+9}{9}$
  \end{compactitem}
\end{cor}

@@

As mentioned in \cref{Introduction}, \citet{DEJGMM21,EJM} prove results about adjacency labelling schemes for planar graphs. Stated in terms of universal graphs, their main theorem is interpreted as follows:

@@

\begin{cor}\label{universal_cor}
	\label{UniversalUniversal}
	For every fixed graph $X$ and all fixed integers $d,\delta,\Delta,g,k>0$, and every integer $n>0$, there exists a graph $U_n$ with $n^{1+o(1)}$ vertices and edges such that $U_n$ contains all of the following graphs as induced subgraphs:
	\begin{compactitem}
		\item every $n$-vertex $(g,k)$-planar graph;
		\item every $n$-vertex $(g,d)$-map graph;
		\item every $n$-vertex $(g,\delta)$-string graph;
		\item every $n$-vertex graph $G^k$ where $G$ is $X$-minor-free and has maximum degree at most $\Delta$;
		\item every $k$-nearest neighbour graph of $n$ points in $\R^2$.
	\end{compactitem}
\end{cor}

>>>>

\begin{table}[h]
  \centering{
    \begin{tabular}{lccl}
      \textbf{Graph class} & \boldmath $\tw(H)$ & \boldmath $\ell$ & \textbf{Reference} \\ \hline
      planar & $3^*$ & $3$ & \cite{DJMMUW20} \\
      planar & $6^*$ & $1$ & \cite{UWY} \\
      genus $g$ & $3^*$ & $\max\{2g,3\}$ & \cite{DHHW} \\
%      genus $g$ & $2g+6$ & 1 & \cite{UWY} \\
      apex-minor-free & $O(1)$ & $O(1)$ & \cite{DJMMUW20} \\
      % B-minor-free & $O(1)$ & $O(1)$ & \cite{DJMMUW20} \\
      $k$-planar & $\binom{k+4}{3}-1$ & $18k^2 + 48k + 30$ & \cref{kPlanarProduct} \\
      $(g,k)$-planar & $\binom{k+4}{3}-1$ & $\max\{2g,3\}(6k^2+16k+10)$ & \cref{gkPlanarProduct} \\
      $(g,\delta)$-string & $\binom{\delta+4}{3}-1$ & $\max\{2g,3\}(\delta^4 + 4\delta^3 + 9\delta^2 + 10\delta+4)$ & \cref{StringProduct} \\
      $k$-nearest-neighbour & $O(k^6)$ & $O(k^4)$ & \cref{nn_product_structure} \\
      $d$-framed & $3^*$ & $d+ 3\floor{d/2} - 3$ & \cref{d_framed_product_stucture} \\
      $1$-planar & $3^*$ & $7$ & \cref{1_planar_product} \\
      $d$-map & $3^*$ & $d+ 3\floor{d/2} - 3$ & \cref{dMapProduct} \\
      $(g,d)$-map & $9$ & $\max\{2g,3\}\,(7d^2 -21d)$ & \cref{gdMapProduct} \\ \hline 
      \multicolumn{4}{l}{$^*$ these bounds also apply to the simple treewidth of $H$.}
    \end{tabular}
  }
  \caption{Known product structure theorems (of the form $G\subseteq H\boxtimes P\boxtimes K_\ell$).}
  \label{summary_table}
\end{table}

\note{DW}{why is the $2g+6$ commented out from the table?}

@@

\note{DW}{Suggested section name ``Graphs Derived From Framed Graphs'' or just 
``Framed Graphs''}

@@

Every $d$-framed multigraph can be described by a $(2,d(d-3)/2)$-shortcut system applied to a plane multigraph by adding a vertex inside each face $F$ of $G_0$ adjacent to the vertices of $F$ and creating a shortcut between each non-consecutive pair of vertices of $F$. Thus \cref{ShortcutProduct} is applicable. However, for framed graphs we prove the following stronger result, where the treewidth bound on $H$ is best possible.\footnote{A multigraph $G$ is \defin{contained} in a graph $X$ if the simple graph underlying $G$ is contained in $X$.}

@@

	The proof is by induction on the number of inner faces of $N_0$.  A particular easy case occurs when $N_0$ has no inner vertices, i.e., $V(N_0)=V(F)$. In this case we can simply take
	\note{DW}{Suggested revision, ``In the base case, $N_0$ has only one inner face, and we can simply take ...} 
  % \note{PM}{I debated what to put here since, even when $N'$ has only one face, the rules below work for picking $\tau$.  The \emph{real} base case occurs when $N'$ has zero faces, but that's not something that ever happens since we only ever apply induction on the inner faces of $M$ (except $\tau$).  I settled on this ``base case'' since it's clear that, in this case, the partition $\mathcal{P}$ is already decided.}
  \note{PM}{The longwinded version is that induction is on $(x,y)$ where $x$ is the number of inner vertices of $N_0$ and $y$ is the number of inner faces of $N_0$.  The base cases are $(0,y)$ for any $y$.  Unfortunately, we can't just induct on $x$ because it doesn't always decrease.} \note{DW}{We need to make it clear what we are doing induction on. 
I suggest we write, ``we proceed by  
induction on $(x,y)$ where $x$ is the number of inner vertices of $N_0$ and $y$ is the number of inner faces of $N_0$. ''}
	$\PP:=\{P_1,\ldots,P_k\}$ and verify that the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements (i)--(iii):
	\begin{compactenum}[(i)]
		\item By definition, each of $P_1,\ldots,P_k$ is a part of $\mathcal{P}$.
		\item Assumption 4(b) ensures, for each $x\in V(H)$, the existence of $X_x$ and $Y_x$ such that $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each integer $j\ge 0$.
		\item The trivial tree decomposition of $H$ that has a single bag containing $\{x_1,\ldots,x_k\}$ is $3$-simple and, if $k=3$, has exactly one bag that contains $\{x_1,\ldots,x_3\}$.
	\end{compactenum}

@@

\begin{proof}
% \note{PM}{I fixed some issues here that come from the fact that some vertices of $G_0$ do not contribute faces to $G_1$.  This happens when two nations and two lakes meet at a vertex $x$.  It's not an issue but it means we can't write ``the face of $G_1$ that corresponds to $x$.''} \note{DW}{Good}
Let $G_0$ be a graph embedded in the plane, with each face labelled a nation or a face, and where each vertex of $G_0$ is incident with at most $d$ nations. If two lakes have a common edge in their boundaries, then we may delete the edges and merge the lakes \note{DW}{Delete the previous sentence? This sentence is incompatible with ``in the clockwise or anticlockwise order of faces around $x$, all the faces between $A$ and $B$ are lakes'' below.}. Now assume that no two lakes have a common edge in their boundaries.
Let $G$ be the corresponding map graph.
Let $G_1$ be the graph with $V(G_1):=V(G)$, which is the set of nations, and whose edge set we now define.  Consider vertices $v$ and $w$ of $G_1$.
Let $A$ and $B$ be the nation faces of $G_0$ corresponding to $v$ and $w$.
Then $vw\in E(G_1)$ if $A$ and $B$ have a common edge in their boundaries,
or there is a vertex $x$ of $G_0$ in the boundary of $A$ and $B$, and in the clockwise or anticlockwise order of faces around $x$, all the faces between $A$ and $B$ are lakes.
% Then $G_1$ has Euler genus at most $g$.
Each face of $G_1$ corresponds to a vertex of $G_0$ or a lake of $G_0$.
Triangulate each face of $G_1$ corresponding to a lake of $G_0$.

For each face $F$ of $G_1$ that corresponds to a vertex $x$ of $G_0$, the size of $F$ \note{DW}{Is ``size'' defined? Do we mean length of the facial cycle? } equals the number of nations incident to $x$, which is at most $d$. So every face of $G_1$ has size at most $d$.  Let $\widehat{G}_1$ be the $d$-framed graph whose frame is $G_1$.  \note{DW}{I like this notation. Should we use it throughout?}

To complete the proof, we now show that $\widehat{G}_1$ contains every edge of $G$.
If $vw\in E(G)$ then the nation faces of $G_0$ corresponding to $v$ and $w$ have a vertex $x$ of $G_0$ in common.   If some face of $G_1$ corresponds to $x$, then $v$ and $w$ are on a common face of $G_1$ so $vw\in E(\widehat{G}_1)$.  If no face of $G_1$ corresponds to $x$\footnote{This can occur when $x$ is on the boundary of exactly two nation faces and two lake faces.}, then the only two nation faces of $G_0$ with $x$ on their boundary are those corresponding to $v$ and $w$, in which case $vw\in E(G_1)\subseteq E(\widehat{G}_1)$.  Therefore $G\subseteq\widehat{G}_1$.
\end{proof}

@@

As discussed in the introduction, product structure has been used to resolve a number of problems on planar graphs.  In most cases, these results hold for any graph class with product structure. In this section, we survey some of the consequences of this for the graph classes considered in the previous two sections.

@@

For an integer $k\geq 0$, a \defin{$k$-queue layout} of a graph $G$ consists of a linear ordering $\preceq$ of $V(G)$ and a partition $\{E_1,E_2,\dots,E_k\}$ of $E(G)$, such that for $i\in\{1,2,\dots,k\}$, no two edges in $E_i$ are nested with respect to $\preceq$. That is, it is not the case that $v\prec x \prec y \prec w$ for edges $vw,xy\in E_i$. The \defin{queue-number} of a graph $G$, denoted by \defin{$\qn(G)$}, is the minimum integer $k$ such that $G$ has a $k$-queue layout. Queue-number was introduced by \citet{HLR92}, who famously conjectured that planar graphs have bounded queue-number. \citet{DJMMUW20} recently proved this conjecture using \cref{PlanarProduct} and the following lemma. Indeed, resolving this question was the motivation for the development of \cref{PlanarProduct}.

@@

Since all the product structure theorems presented thus far upper bound the treewidth of $H$, \cref{qn} immediately implies bounds on the queue number of all graphs in these classes.  An improvement can be made for $1$-planar and $d$-map graphs since then $H$ is planar with treewidth at most $3$. \citet{ABGKP20} proved that every planar graph with treewidth at most $3$ has queue-number at most $5$. Thus the graph $H$ in \cref{1_planar_product,dMapProduct} has queue-number at most $5$.
% , which with \cref{qn} implies:
% 
% \begin{cor}
% \label{1PlanarQueue}
% Every 1-planar graph has queue-number at most 
% $3 \times 7 \times 5 + \floor{\tfrac{3}{2} \times 7} = 115$.
% \end{cor}
% 
% \begin{cor}
% \label{dMapQueue}
% Every $d$-map graph has queue-number at most 
% $\floor{ \frac{33}{2} (d+3\floor{d/2}-3) }$.
% \note{DW}{I have updated this bound using the better qn bound for planar 3-trees. Note that with $d=3$ we get the original bound of 49.}
% \end{cor}
% 
% 
% 
% \cref{qn,gkPlanarProduct} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^3)}$.
% \note{DW}{I have changed $g 2^{O(k^4)}$ to $g 2^{O(k^3)}$, okay?}
% \note{OM}{Ok}
% \note{DW}{I have changed $g>2^{k^3}$ to $g>2^{k^2}$, okay?}.
% \note{PM}{Ok}
The following corollary summarizes all of these results.
\begin{cor}\label{q_cor}
  The following bounds hold for the queue number of any graph from each of the following classes:
  \begin{compactitem}
    \item $k$-planar: $2^{O(k^3)}$
    \item $(g,k)$-planar: $\max\{1,g\}\cdot 2^{O(k^3)}$
    \item $(g,\delta)$-string: $\max\{1,g\}\cdot 2^{O(\delta^3)}$
    \item $1$-planar: $115$ 
    \item $d$-map: $\floor{ \frac{33}{2} (d+3\floor{d/2}-3) }$ 
    \item $(g,d)$-map: $\max\{1,g\}\cdot 2^{O(d)}$
    \item $k$-nearest-neighbour: $2^{O(k^6)}$
  \end{compactitem}
\end{cor}
%\note{PM}{For uniformity, I got rid of the two corollaries on $d$-map and $1$-planar graphs and kept them in \cref{q_cor}.  The text is still commented out in the source in case someone feels strongly about keeping them.}

@@

Combining \cref{non-repetitive} with our results on product structure, we obtain the following corollary:
\begin{cor}\label{nonrepetitive_cor}
  The following bounds hold for the nonrepetitive chromatic number of any graph from each of the following classes:
  \begin{compactitem}
    \item $k$-planar: $(18k^2+48k+30)\cdot 4^{\binom{k+4}{3}}$
    \item $(g,k)$-planar: $\max\{2g,3\}\cdot(6k^2+16k+10) 4^{\binom{k+4}{3}}$
    % \item $(g,\delta)$-string: $g4^{O(\delta^3)}$
    \item $1$-planar: $1792$
    \item $d$-map: $256(d+3\floor{d/2}-3)$
    % \item $(g,d)$-map: $g4^{O(d)}$
    % \item $k$-nearest-neighbour: $4^{O(k^6)}$
    \item $(g,d)$-map graph: $(7d^2-21d)\cdot 4^{10}$
  \end{compactitem}
\end{cor}

@@

\begin{cor}\label{p_centered_cor}
  The following bounds hold for the $p$-centered chromatic number of any graph from each of the following classes:
  \begin{compactitem}
    \item $k$-planar: $(18k^2+48k+30)\cdot 4^{\binom{k+4}{3}}$
    \item $(g,k)$-planar: $\max\{2g,3\}\cdot(6k^2+16k+10)(p+1)\binom{p+\binom{k+4}{3}-1}{\binom{k+4}{3}-1}$
    % \item $(g,\delta)$-string: $g4^{O(\delta^3)}$
    \item $1$-planar: $7(p+1)^2(p\ceil{\log(p+1)}+2p+1)$
    \item $d$-map: $(d+3\floor{d/2}-3)\cdot (p+1)^2(p\ceil{\log(p+1)}+2p+1)$
    % \item $(g,d)$-map: $g4^{O(d)}$
    % \item $k$-nearest-neighbour: $4^{O(k^6)}$
    \item $(g,d)$-map: $\max\{2g,3\}\cdot 7d(d-3)(p+1)\cdot \binom{p+9}{9}$
  \end{compactitem}
\end{cor}

@@

As mentioned in \cref{Introduction}, \citet{DEJGMM21} and \citet{EJM} prove results about adjacency labelling schemes for planar graphs. Stated in terms of universal graphs, their main theorem is interpreted as follows:

@@

\begin{cor}\label{universal_cor}
	\label{UniversalUniversal}
	For every fixed graph $X$ and all fixed integers $d,\delta,\Delta,g,k>0$, and every integer $n>0$, there exists a graph $U_n$ with $n^{1+o(1)}$ vertices and edges such that $U_n$ contains the following graphs as induced subgraphs:
	\begin{compactitem}
		\item every $n$-vertex $(g,k)$-planar graph;
		\item every $n$-vertex $(g,d)$-map graph;
		\item every $n$-vertex $(g,\delta)$-string graph;
		\item every $n$-vertex graph $G^k$ where $G$ is $X$-minor-free and has maximum degree at most $\Delta$;
		\item every $k$-nearest neighbour graph of $n$ points in $\R^2$.
	\end{compactitem}
\end{cor}
